Liveness-Based Garbage Collection for Lazy Languages

### 2015-06-03 paper/paper.tex
liveness

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
\cmt{{
\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
section.
%
\author{
% 1st. author
\alignauthor K. Prasanna Kumar \\
       \affaddr{IIT Bombay,}\\
       \affaddr{Mumbai 400076, India}
       \email{prasanna@cse.iitb.ac.in}
% 2nd. author
\alignauthor Amey Karkare\\
       \affaddr{IIT Kanpur}\\
       \affaddr{Kanpur 208016, India}
       \email{karkare@cse.iitk.ac.in}
% 3rd. author
\alignauthor Amitabha Sanyal \\
       \affaddr{IIT Bombay}\\
       \affaddr{Mumbai 400076, India}\\
       \email{as@cse.iitb.ac.in}
}
}}
\authorinfo{Double Blind Review}{}{}
\maketitle

@@

\subsubsection{Semantics}
We now give  a small-step semantics for our  language.
We first specify the domains used by the semantics:
\[
\begin{array}{rlcl@{\hspace{0.5em}}l}
\rho: & \mathit{Env} &=&\mathit{Var} \rightarrow \mathit{Loc} &
\mbox{-- Environment} \\
v:   & \mathit{Val} &=& \mathbb{N} + \{\NIL\} + \mathit{Data \times
  Data}& \mbox{-- Values}\\
c:   & \mathit{Clo} &=& \mathit{(Exp \times Env)}& \mbox{--
  Closures}\\
d: & \mathit{Data} &=&\mathit{Val} + \mathit{Clo} & \mbox{-- Values \&
  Closures} \\
\heap: & \mathit{Heap} & =&\mathit{Loc} \rightarrow Data & \mbox{--
Heap}
\end{array}
\]

@@

A  notion  related  to  liveness  is  {\em  demand}.   While  liveness
represents the future use of variables, a demand represents the future
use of the value of an  expression. The demand on an expression $e$ is
again a  set of  access paths---that subset  of $\{\acar,\acdr\}^\ast$
which the context of $e$ may explore of $e$'s result.  To see the need
for  demands,  consider   the  let  expression  $\pi:(\LET~x\leftarrow
(\CDR~y)~\IN~  \pi':\SRETURN~x)$.   Assume that  the  context of  this
expression produces  the liveness $[x \mapsto  \lbrace \epsilon, \acar
  \rbrace]$ at $\pi'$.  Due to the \LET\ definition which binds $x$ to
$(\CDR~y)$, the  liveness of $x$ at  $\pi'$ now becomes  the demand on
$(\CDR~y)$.    This,  in   turn,  generates   the   liveness  $\lbrace
y.\epsilon, y.\acdr, y.\acdr\acar \rbrace$ at $\pi$. These are the set
of $y$-rooted  accesses required  to explore $\lbrace  \epsilon, \acar
\rbrace$  paths of  the result  of  $(\CDR~y)$\footnote{The classical
  analogue  is strong  liveness analysis,  where  $y$ and  $z$ have  a
  liveness of  $\lbrace\epsilon\rbrace$ at  the entry $n:  x:=y+z$, if
  and only if  $x$ has a liveness of  $\lbrace\epsilon\rbrace$ at exit
  of $n$.   In our terminology, the  liveness of $x$ at  exit from $n$
  becomes the demand on $y+z$.   This, in turn, generates the liveness
  $\lbrace y.\epsilon, z.\epsilon \rbrace$ at the entry of $n$.}.
%% Exploring $\lbrace \epsilon, \acar \rbrace$ of the
%% result of $(\CDR~y)$ first requires access using $y$ ( and the \CDR\
%% field of the cons cell pointed by
%% $y$.

@@

Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The function $\mathit{ref}$,  when given an application $s$ and
a demand $\sigma$, returns  the incremental liveness generated for the
free  variables   of  $s$  due  to  the   application.   The  function
$\mathcal{L}$  uses   $\mathit{ref}$  to  propagate   liveness  across
expressions.

 In each  rule defining $\mathit{ref}$, the  null demand ($\emptyset$)
 generates no  liveness.  Notice that this is  different from liveness
 of  eager  languages,  where,   for  instance,  (\CDR~$x$)  would  be
 evaluated  even for  the null  demand. This  would require  an access
 using $x$ resulting in  $x.\epsilon$ being live.  A non-null demand of
 $\sigma$ on (\CDR~$x$), is transformed to the liveness $\{x.\epsilon,
 x.\acdr\sigma\}$.  In an opposite  sense, the demand of $\acdr\sigma$
 on  (\CONS~$y$~$z$) is  transformed to  the demand  $\sigma$  on $z$.
 Note that \CONS\ does not,  by itself, dereference its arguments. The
 rules for (\PRIM~x~y) and  (\NULLQ~x) are on similar lines. Constants
 generate no liveness.

For  the  case  of a  user-function  call  we  use a  third  parameter
$\Lfonly$  that  represents the  summaries  of  all  functions in  the
program.   $\Lfonly_{\mathit f}$  (the  component of  $\Lfonly$ for  a
specific  function $f$),  expresses  how the  demand  $\sigma$ on  the
result  of a  call to  $f$  is transformed  into the  liveness of  its
parameters at the  beginning of the call.  $\Lfonly$  is determined by
the judgement  form $\mathit{Prog} \len \Lfonly$  using inference rule
({\sc live-define}).  This rule  describes the fixed-point property to
be satisfied  by $\Lfonly$, namely, the  demand transformation assumed
for each  function in  the program  should be the  same as  the demand
transformation  calculated from  their  bodies.  As  we  shall see  in
Section~\ref{sec:grammar-formulation},  we  convert  the rule  into  a
grammar and  the least  solution satisfying the  rule is  the language
generated by this  grammar.  The least solution also  ensures the safe
collection of the greatest amount of garbage.

We next  describe the function $\mathcal{L}$  that propagates liveness
across expressions.  Note that,  unlike eager languages, the syntactic
occurrence  of an expression  in a  program denotes  the point  of
creation of its closure, not necessarily its evaluation.
To see the  consequences of this, consider
Figure~\cite{fig:mot-example2} which shows an analysis  of the body of
a function $\length$ with a demand $\sigma$.  Clearly, this is also the
demand on the return value $z$ giving the liveness at the program point
$\pi_8$ as $\{\pz.\sigma\}$.  The liveness at $\pi_8$ is propagated
backwards as in traditional liveness analysis.

@@

        \small
\begin{eqnarray*}
        \mathcal{L}(e_{\length},\sigma,\Lfonly) &=&
              [\;ep_1 \mapsto \lbrace \px.\sigma_{\length},
\xl.\cbar\sigma_{\length} \rbrace ,~ep_2 \mapsto \lbrace \rbrace,\\
 &&          \;\;  ep_3 \mapsto \lbrace \pz.\sigma_{\length},
                \py.\cbar\sigma_{\length},
\pu.\Lf{\length}{\mbox{1}}{\cbar\sigma_{\length}},\\
&&            \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
\xl.\acdr\Lf{\length}{\mbox{1}}{\cbar\sigma_{\length}}
                \rbrace~\rbrack \\
            \cal{P_{\mbox{\tiny \length}}} &=&
                [\;\pi_1 \mapsto \lbrace ep_1, ep_2, ep_3 \rbrace,  \
                 \pi_2 \mapsto \lbrace ep_1, ep_2, ep_3
                  \rbrace, \\
&&\;\;  \pi_3 \mapsto \lbrace ep_2
                  \rbrace ,  \pi_4 \mapsto \lbrace ep_2
                  \rbrace, \pi_5 \mapsto \lbrace  ep_3 \rbrace, \\
 &&  \;\;                        \pi_6 \mapsto \lbrace ep_3 \rbrace,
\pi_7 \mapsto
                  \lbrace ep_3\rbrace, \pi_8 \mapsto \lbrace ep_3
\rbrace~\rbrack\\
              \mathcal{L}(e_{\main},\sigma,\Lfonly) &=&
                [\;
                  ep_4 \mapsto \lbrace \pw.\sigma_{\mathit{\!all}},
\pb.\Lf{\length}{\mbox{1}}{\sigma_{\mathit{\!all}}}, \ldots\rbrace
                  ]\\
              \cal{P_{\mbox{\tiny \main}}} &=&
                  \lbrack~\pi_6 \mapsto \lbrace ep_4 \rbrace,  \
                   \pi_7 \mapsto \lbrace ep_4 \rbrace,  \
                  \pi_8 \mapsto \lbrace ep_4 \rbrace~\rbrack
\end{eqnarray*}
    \end{minipage} \\
(a)&
(b)
    %% \end{ tabular}}
  \end{tabular}
%\end{picture}
\kern -3ex
%\vspace{0.5in}
\caption{(a) Example program and (b) its liveness
maps.}\label{fig:mot-example2}
\end{figure*}

@@

 The liveness  of {\LET} has  an elegant formulation.   Recollect that
 our language permits lazy {\LET}s whose definitions can be recursive,
 i.e.  in the expression $\LET~x \leftarrow~s~\IN~e$, $x$ can occur in
 $s$.   We  model  the  definition  $x  \leftarrow~s$  as  a  possibly
 recursive function (named $\mathit{f}$ in the rule) which captures an
 arbitrary number of ``unrollings'' of the definition\footnote{The $*$
   in  the definition  of ${\mathit  f}$  represents non-deterministic
   choice,  but  can  be  treated   as  a  constant  for  purposes  of
   analysis.}.   We pass to  the demand  transformer $\Lfonly_{\mathit
   f}$ of this function the demand  arising out of the liveness of $x$
 at the beginning of $e$. This  gives the liveness of the variables of
 $s$.

@@

Section~\ref{sec:computing}   shows   how   the  demand   transformers
$\Lfonly$  for  a  program  (representing  a  fully  context-sensitive
analysis) can be safely approximated  by a {\em procedure summary} for
each function  (unifying the contexts  in the style of  0-CFA)\@.  The
summary consists  of a pair of  a single demand and,  for this demand,
the  corresponding  tuple  of   demands  the  function  makes  on  its
arguments.

@@

\section{Computing liveness and its encoding as a
table}\label{sec:computing}
Section~\ref{sec:liveness} gave  a context-sensitive liveness analysis
and  {\color  {red}  proved  it  correct  with  reference  to  a  {\em
    minefield} semantics}.  In  particular, $\mathcal{L}$ and \pmap{f}
together described a liveness set for each program point in a function
in terms of  a given $\sigma$ and \Lfonly.  However,  we still have to
describe how to obtain demand transformers \Lfonly\ from the rule {\sc
  live-define} and how to compute the specific demand $\sigma$ on each
function.

To  answer these  questions, it  is convenient  to cast  the equations
arising out of the $\mathit{ref}$  and $\mathcal{L}$ as the rules of a
grammar.   But first  we modify  the rules  themselves to  a different
form.

\subsubsection{Modifying the liveness rules.}

@@

@@

{\color{red} \medskip \hrule \hrule \hrule  \medskip }

@@

\section{Generating Liveness Equations}
The  program  in  Figure~\ref{fig:mot-example2}  serves as  a  running
example.   Given a  function $\mathit{f}$,  we now  describe  how to
generate    equations    for     the    demand    transformation
\Lfonly$_\mathit{f}$.   Starting with a  symbolic demand  $\sigma$, we
determine  \Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.  In  particular, we
consider $\Lanv{1}{x_i}$,  the  liveness of    the
$i^{\mbox{\footnotesize  th}}$
parameter $x_i$ at the  program point at the  beginning of
$e_{\mathit{f}}$ (assumed  to be  $\pi_1$). By the rule {\sc
  live-define}, this should be the
same as $\Lf{f}{i}{\sigma}$. Applying this to  \length, we have:
\begin{eqnarray*}
&& \Lanv{1}{\xl} = \cbar\sigma \cup \acdr\Lf{\length}{1}{\cbar\sigma}
\end{eqnarray*}
and  the only equation defining \Lfone{\length} is:
  \begin{eqnarray*}
   && \Lf{\length}{1}{\sigma}
    =  \cbar\sigma \cup \acdr\Lf{\length}{1}{\cbar\sigma}
   \end{eqnarray*}

>>>>

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
\cmt{{
\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
section.
%
\author{
% 1st. author
\alignauthor K. Prasanna Kumar \\
       \affaddr{IIT Bombay,}\\
       \affaddr{Mumbai 400076, India}
       \email{prasanna@cse.iitb.ac.in}
% 2nd. author
\alignauthor Amey Karkare\\
       \affaddr{IIT Kanpur}\\
       \affaddr{Kanpur 208016, India}
       \email{karkare@cse.iitk.ac.in}
% 3rd. author
\alignauthor Amitabha Sanyal \\
       \affaddr{IIT Bombay}\\
       \affaddr{Mumbai 400076, India}\\
       \email{as@cse.iitb.ac.in}
}
}}
\authorinfo{Double Blind Review}{}{}
\maketitle

@@

\subsection{Semantics}
We now give  a small-step semantics for our  language.
We first specify the domains used by the semantics:
\[
\begin{array}{rlcl@{\hspace{0.5em}}l}
\rho: & \mathit{Env} &=&\mathit{Var} \rightarrow \mathit{Loc} &
\mbox{-- Environment} \\
v:   & \mathit{Val} &=& \mathbb{N} + \{\NIL\} + \mathit{Data \times
  Data}& \mbox{-- Values}\\
c:   & \mathit{Clo} &=& \mathit{(Exp \times Env)}& \mbox{--
  Closures}\\
d: & \mathit{Data} &=&\mathit{Val} + \mathit{Clo} & \mbox{-- Values \&
  Closures} \\
\heap: & \mathit{Heap} & =&\mathit{Loc} \rightarrow Data & \mbox{--
Heap}
\end{array}
\]

@@

A  notion  related  to  liveness  is  {\em  demand}.   While  liveness
represents the future use of variables, a demand represents the future
use of the value of an  expression. The demand on an expression $e$ is
again a  set of  access paths---that subset  of $\{\acar,\acdr\}^\ast$
which the context of $e$ may explore of $e$'s result.  To see the need
for  demands,  consider   the  let  expression  $\pi:(\LET~x\leftarrow
(\CDR~y)~\IN~  \pi':\SRETURN~x)$.   Assume that  the  context of  this
expression produces  the liveness $[x \mapsto  \lbrace \epsilon, \acar
  \rbrace]$ at $\pi'$.  Due to the \LET\ definition which binds $x$ to
$(\CDR~y)$, the  liveness of $x$ at  $\pi'$ now becomes  the demand on
$(\CDR~y)$.    This,  in   turn,  generates   the   liveness  $\lbrace
y.\epsilon, y.\acdr, y.\acdr\acar \rbrace$ at $\pi$. These are the set
of $y$-rooted  accesses required  to explore $\lbrace  \epsilon, \acar
\rbrace$  paths   of  the  result   of  $(\CDR~y)$\footnote{Reproduced
  from~\cite{asati14lgc}: In  classical strong liveness  analysis, $y$
  and $z$  are live at the  entry $n: x:=y+z$,  if and only if  $x$ is
  live  at exit  of $n$.   In our  terminology, the  liveness $\lbrace
  \epsilon\rbrace $ of $x$ at the  exit from $n$ becomes the demand on
  $y+z$, and this, in turn generates the liveness $\lbrace y.\epsilon,
  z.\epsilon \rbrace$ at the entry of $n$.}.
%% Exploring $\lbrace \epsilon, \acar \rbrace$ of the
%% result of $(\CDR~y)$ first requires access using $y$ ( and the \CDR\
%% field of the cons cell pointed by
%% $y$.

@@

Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The function $\mathit{ref}$,  takes  an application $s$ and
a demand $\sigma$, and returns  the incremental liveness generated for the
free  variables   of  $s$  due  to  the   application.   The  function
$\mathcal{L}$  uses   $\mathit{ref}$  to  propagate   liveness  across
expressions.

 Since  in a  lazy language,  an  expression is  not evaluated  unless
 required, the null demand ($\emptyset$) does not generate liveness in
 any  of  the  rules  defining  $\mathit{ref}$  or  $\mathcal{L}$.   A
 non-null  demand of  $\sigma$ on  (\CDR~$x$), is  transformed  to the
 liveness $\{x.\epsilon, x.\acdr\sigma\}$.   In an opposite sense, the
 demand  of $\acdr\sigma$  on  (\CONS~$y$~$z$) is  transformed to  the
 demand  $\sigma$ on  $z$.   Note  that \CONS\  does  not, by  itself,
 dereference its  arguments. The rules for  (\PRIM~x~y) and (\NULLQ~x)
 are similar. Constants do not generate any liveness.

For  the  case  of a  user-function  call  we  use a  third  parameter
$\Lfonly$  that  represents the  summaries  of  all  functions in  the
program.   $\Lfonly_{\mathit f}$  (the  component of  $\Lfonly$ for  a
specific function $f$), expresses how the demand $\sigma$ on a call to
$f$  is  transformed  into  the  liveness of  its  parameters  at  the
beginning of the call.  $\Lfonly$  is determined by the judgement form
$\mathit{Prog} \len \Lfonly$ using inference rule ({\sc live-define}).
This  rule  describes the  fixed-point  property  to  be satisfied  by
$\Lfonly$, namely, the demand transformation assumed for each function
in  the  program should  be  the  same  as the  demand  transformation
calculated    from    its    body.      As    we    shall    see    in
Section~\ref{sec:grammar-formulation},  we  convert  the rule  into  a
grammar  and the  the  language  generated by  this  grammar is  least
solution satisfying  the rule. We  prefer the least solution  since it
ensures the safe collection of the greatest amount of garbage.

We next  describe the function $\mathcal{L}$  that propagates liveness
across expressions.  Note that,  unlike eager languages, the syntactic
occurrence  of an expression  in a  program denotes  the point  of
creation of its closure, not necessarily its evaluation.
To see the  consequences of this, consider
Figure~\ref{fig:mot-example2} which shows an analysis  of the body of
a function $\length$ with a demand $\sigma$.  Clearly, this is also the
demand on the return value $z$ giving the liveness at the program point
$\pi_8$ as $\{\pz.\sigma\}$.  The liveness at $\pi_8$ is propagated
backwards as in traditional liveness analysis.

@@

@@

 The liveness  of {\LET} has  an elegant formulation.   Recollect that
 our language permits lazy {\LET}s whose definitions can be recursive,
 i.e.  in the expression $\LET~x \leftarrow~s~\IN~e$, $x$ can occur in
 $s$.   We  model  the  definition  $x  \leftarrow~s$  as  a  possibly
 recursive function (named $\mathit{f}$ in the rule) which captures an
 arbitrary number of ``unrollings'' of the definition\footnote{The $*$
   in  the definition  of ${\mathit  f}$  represents non-deterministic
   choice.  During   liveness  it  can  be  assumed   to  generate  no
   liveness.}.   We pass to  the demand  transformer $\Lfonly_{\mathit
   f}$ of this function the demand  arising out of the liveness of $x$
 at the beginning of $e$. This  gives the liveness of the variables of
 $s$.

@@

Section~\ref{sec:computing}   shows   how   the  demand   transformers
$\Lfonly$  for  a  program  (representing  a  fully  context-sensitive
analysis) can be safely approximated by a {\em procedure summary}. for
each function.   The summary  is in the  form of a  demand transformer
that maps a demand on a call to the function to demands on each of its
arguments.

@@

\section{Computing liveness and its encoding as DFA}\label{sec:computing}
Section~\ref{sec:liveness} gave  a context-sensitive liveness analysis
and  {\color  {red}  proved  it  correct  with  reference  to  a  {\em
    minefield} semantics}.  In  particular, $\mathcal{L}$ and \pmap{f}
together described a liveness set for each program point in a function
in terms of  a given $\sigma$ and \Lfonly.  However,  we still have to
describe how to obtain demand transformers \Lfonly\ from the rule {\sc
  live-define} and how to compute the specific demand $\sigma$ on each
function.

To  answer these  questions, it  is convenient  to cast  the equations
arising out of $\mathit{ref}$  and $\mathcal{L}$ as the rules of a
grammar.   But first  we modify  the rules  themselves to  a different
form.

\subsection{Modifying the liveness rules.}

@@

        \small
\begin{eqnarray*}
        \mathcal{L}(e_{\length},\sigma,\Lfonly) &=&
              [\;ep_1 \mapsto \lbrace \px.\sigma_{\length},
\xl.\cbar\sigma_{\length} \rbrace ,~ep_2 \mapsto \lbrace \rbrace,\\
 &&          \;\;  ep_3 \mapsto \lbrace \pz.\sigma_{\length},
                \py.\cbar\sigma_{\length},
\pu.\Lf{\length}{\mbox{1}}{\cbar\sigma_{\length}},\\
&&            \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
\xl.\acdr\Lf{\length}{\mbox{1}}{\cbar\sigma_{\length}}
                \rbrace~\rbrack \\
            \cal{P_{\mbox{\tiny \length}}} &=&
                [\;\pi_1 \mapsto \lbrace ep_1, ep_2, ep_3 \rbrace,  \
                 \pi_2 \mapsto \lbrace ep_1, ep_2, ep_3
                  \rbrace, \\
&&\;\;  \pi_3 \mapsto \lbrace ep_2
                  \rbrace ,  \pi_4 \mapsto \lbrace ep_2
                  \rbrace, \pi_5 \mapsto \lbrace  ep_3 \rbrace, \\
 &&  \;\;                        \pi_6 \mapsto \lbrace ep_3 \rbrace,
\pi_7 \mapsto
                  \lbrace ep_3\rbrace, \pi_8 \mapsto \lbrace ep_3
\rbrace~\rbrack\\
              \mathcal{L}(e_{\main},\sigma,\Lfonly) &=&
                [\;
                  ep_4 \mapsto \lbrace \pw.\sigma_{\mathit{\!all}},
\pb.\Lf{\length}{\mbox{1}}{\sigma_{\mathit{\!all}}}, \ldots\rbrace
                  ]\\
              \cal{P_{\mbox{\tiny \main}}} &=&
                  \lbrack~\pi_6 \mapsto \lbrace ep_4 \rbrace,  \
                   \pi_7 \mapsto \lbrace ep_4 \rbrace,  \
                  \pi_8 \mapsto \lbrace ep_4 \rbrace~\rbrack
\end{eqnarray*}
    \end{minipage} \\
(a)&
(b)
    %% \end{ tabular}}
  \end{tabular}
%\end{picture}
\kern -3ex
%\vspace{0.5in}
\caption{(a) Example program and (b) its liveness
maps.}\label{fig:mot-example2}
\end{figure*}
%================================================================

@@

%Thick red line 
%{\color{red} \medskip \hrule \hrule \hrule  \medskip }

@@

\subsection{Generating Liveness Equations}
The  program  in  Figure~\ref{fig:mot-example2}  serves as  a  running
example.   Given a  function $\mathit{f}$,  we now  describe  how to
generate    equations    for     the    demand    transformation
\Lfonly$_\mathit{f}$.   Starting with a  symbolic demand  $\sigma$, we
determine  \Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.  In  particular, we
consider $\Lanv{1}{x_i}$,  the  liveness of    the
$i^{\mbox{\footnotesize  th}}$
parameter $x_i$ at the  program point at the  beginning of
$e_{\mathit{f}}$ (assumed  to be  $\pi_1$). By the rule {\sc
  live-define}, this should be the
same as $\Lf{f}{i}{\sigma}$. Applying this to  \length, we have:
\begin{eqnarray*}
&& \Lanv{1}{\xl} = \cbar\sigma \cup \acdr\Lf{\length}{1}{\cbar\sigma}
\end{eqnarray*}
and  the only equation defining \Lfone{\length} is:
  \begin{eqnarray*}
   && \Lf{\length}{1}{\sigma}
    =  \cbar\sigma \cup \acdr\Lf{\length}{1}{\cbar\sigma}
   \end{eqnarray*}
